---
title: "R Skills Assignment"
author: "Nadia Matulich"
format: html
---

```{r}

library(tidyverse)
library(knitr)
library(kableExtra)
library(DT)


knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.align="left", fig.pos="H", digits=2)


```

```{r}
tag      <- "202311081903"
base_url <- "https://github.com/randrescastaneda/pub_data/raw/"
data_url <- paste0(base_url, tag, "/data/Rtest1/")

options(scipen = 999) # to ensure numbers don't print in scientific notation

options(digits=1)

```

# Basic Statistics

## Import Data

```{r}

wdi <- readr::read_rds(paste0(data_url, "wdi_in1.Rds"))

```

## Define Commonly Called on Functions

```{r}

weighted_mean <- function(x, weights) {
  sum(x * weights, na.rm = TRUE) / sum(weights, na.rm = TRUE)
}

weighted_sd <- function(x, weights) {
  mean_value <- weighted_mean(x, weights)
  sqrt(sum(weights * (x - mean_value)^2, na.rm = TRUE) / sum(weights, na.rm = TRUE))
}

```

## Summary Statistics of GDP per Capita by Region

### Functions

```{r}
# Cannot figure out why I get 44 observations in SSA here instead of 42

summary<- function(data,x,weights){
  
  summary <- 
    data %>%
    group_by(date, region) %>%
    summarise(
      n = n(),
      mean = weighted_mean(.data[[x]], .data[[weights]]),
      sd = weighted_sd(.data[[x]], .data[[weights]]),
      median = median(.data[[x]], .data[[weights]], na.rm = TRUE),
      min = min(.data[[x]], na.rm = TRUE), #no extra steps for weighting as per capita already considers pop size
      max = max(.data[[x]], na.rm = TRUE) # weighting not performed, as with min
    ) %>%
    arrange(region, date)
  
  summary
  
}

```

### Printing Results

```{r}

summary(data= wdi %>% 
          select(region, date, country, gdp, pop),
        x="gdp",
        weights="pop") %>% 
  head(10) %>% 
  knitr::kable("html") %>%
  kable_styling("striped", full_width = FALSE) 

```


## Question 2

### Functions

```{r}

aggregating_statistics <- function(data, x, weights) {
  
results <- data %>%
    group_by(date, region) %>%
    summarise(
      mean = weighted_mean(.data[[x]], .data[[weights]]),
      sd = weighted_sd(.data[[x]], .data[[weights]]),
      median = median(.data[[x]], .data[[weights]], na.rm = TRUE),
      min = min(.data[[x]], na.rm = TRUE),
      max = max(.data[[x]], na.rm = TRUE)
    ) %>%
    arrange(region, date) %>%
    pivot_longer(
      cols = starts_with(c("mean", "sd", "median", "min", "max")),
      names_to = "statistic",
      values_to = "value"
    ) 

results

}

```

```{r}

collating_table <- function(data, variables) {
  
results_list <- list()

pop<-
    data %>% 
    group_by(region,date) %>% 
    summarise(pop=sum(pop))
  
  for (variable in variables) {
    results_list[[variable]] <- aggregating_statistics(data, x = variable, weights = "pop")
  }

result_table <- 
  bind_rows(results_list, .id = "variable") %>% 
  left_join(pop, by=c("region","date")) %>% 
  pivot_wider(names_from = variable, values_from = value) 

result_table

}

```

### Printing Results

```{r}

options(digits=5)

collating_table(data= wdi %>% 
                  select(region, date, iso3c, lifeex, pov_intl, gdp, pop), 
                variables= c("gdp", "lifeex", "pov_intl")) %>% 
  select(statistic,region,date,pop,lifeex,gdp,pov_intl) %>% 
  head(10) %>% 
  knitr::kable("html") %>%
  kable_styling("striped", full_width = FALSE) 
  

```

## Question 3 Finding Outliers

This one threw me for a loop for some reason, and when I tried to use Q2, it just wouldn't work out. 

There is an error here somewhere, but please see my proposed solution below:

### Functions
```{r}
outliers<-function(){
 
outliers<-wdi %>% 
  select(date,region,country,gdp,pop,lifeex,gini) %>% 
  pivot_longer(cols = c("gdp", "lifeex", "gini"),
      names_to = "indicator",
      values_to = "value") %>% 
  group_by(date,region,indicator) %>% 
  mutate(mean=weighted_mean(value,pop),
         sd=weighted_sd(value,pop),
         upper=mean+2.5*(sd),
         lower=mean-2.5*(sd),
         outlier=ifelse(value<=upper,
                (ifelse(indicator<=lower,"outlier","inlier")),"outlier"))
outliers

}
```

### Printing Results

```{r}
outliers() %>% 
  filter(outlier=="outlier") %>% 
  head(10) %>% 
  knitr::kable("html") %>%
  kable_styling("striped", full_width = FALSE)
```

### Printing Graph

```{r}

outliers() %>% 
  filter(indicator=="lifeex") %>% 
  ggplot()+
  geom_ribbon(aes(x=date,
                  ymin=mean-2.5*(sd),
                  ymax=mean+2.5*(sd), 
                  xmin=first(date),
                  xmax=last(date)), 
              colour="grey", alpha=0.3)+
  geom_point(aes(x=date,y=value,color=region))+
  geom_line(aes(x=date, y=mean))+
  theme_minimal()


```

# Poverty Measures

## Reading in Data

```{r}

data <- readr::read_rds(paste0(data_url, "svy_sim_in1.Rds")) %>% 
  bind_rows(.id = "year") %>% 
  mutate(year = as.numeric(substring(year, 2)))

```

## Question 4

### Key Functions

```{r}

measuring_poverty<-function(x){

measuring_poverty<-data %>% 
  mutate(poor=ifelse(income<=x,1,0)) %>% 
  group_by(year) %>% 
  summarise(pov_line=x,
            headcount=sum(poor*weight)/sum(weight),
            povgap=(sum(ifelse(income<=x,((x-income)/x)*weight,0)))/sum(weight),
            povseverity=(sum(ifelse(income<=x,(((x-income)/x)^2)*weight,0)))/sum(weight),
            )

measuring_poverty

}
```

### Printing Results

```{r}

poverty_levels <- c(2.15, 3.65, 6.85)

poverty_measures <- data.frame()

for (level in poverty_levels) {
  result <- measuring_poverty(level)
  poverty_measures<- bind_rows(poverty_measures, result)
}

poverty_measures %>% 
  head(10) %>% 
  knitr::kable("html") %>%
  kable_styling("striped", full_width = FALSE) 

```

### Graphing Results

```{r}

poverty_measures %>% 
  mutate(year = as.Date(paste0(year, "-01-01")),
         pov_line=factor(pov_line)) %>% 
  ggplot() +
  geom_point(aes(x=year,y=headcount,colour=pov_line))+
  geom_line(aes(x=year,y=headcount,colour=pov_line))+
  labs(x="Year",
       y="Head Count")+
  theme_minimal()+
  scale_colour_manual(values=c("#00193a", "#002b60", "#66a3ff"))+
  theme(legend.position="bottom", legend.title=element_blank())

```

## Question 5

Hello! Here's my attempt. I've never coded Lorenz curves before. Ithink the maths is right, as my Gini co-efficients work out at the end, but I know I am missing something with the 100 points. I don't think I understand the question well enough. I've got too many rows in my table, and I am assuming the 100 points line is intended to speed up the process. I'm presenting the computationally intensive code, then a sampling based code afterwards. 

Please see attempt below.

### Functions

```{r}

lorenz<-function(){

lorenz <-data%>% 
  group_by(year) %>%
  arrange(income) %>% 
  mutate(cum_income=cumsum(income*weight)/sum(income*weight),
         cum_population=cumsum(weight)/sum(weight)) 

lorenz

}

```

### Printing Results

```{r}

lorenz() %>% 
  head(10) %>% 
  knitr::kable("html") %>%
  kable_styling("striped", full_width = FALSE) 

```

### Graphing Results

```{r}

lorenz() %>% 
  ggplot() +
  geom_line(aes(x = cum_population, 
                y = cum_income, 
                group = year, 
                color = as.factor(year))) +
  labs(x="Cumulative Welfare",
       y="Cumulative Population")+
  theme_minimal()+
  theme(legend.position=c(0.10,0.60), 
        legend.title=element_blank())


```
## Question 5 (Alternative)

### Functions

This is also not quite right, but shows code based on taking a random sample from the population to calculate the Lorenz curve.

```{r}

set.seed=(123)

lorenz_2<-function(){

sample_data <- data %>%
  group_by(year) %>%
  slice_sample(n = 100, replace = FALSE)
    
lorenz <-sample_data%>% 
  group_by(year) %>%
  arrange(income) %>% 
  mutate(cum_income=cumsum(income*weight)/sum(income*weight),
         cum_population=cumsum(weight)/sum(weight)) %>% 
  arrange(cum_population)

lorenz

}

```

### Printing Results

```{r}

lorenz_2() %>% 
  head(10) %>% 
  knitr::kable("html") %>%
  kable_styling("striped", full_width = FALSE) 


```

### Graphing Results

```{r}

lorenz_2() %>% 
  ggplot() +
  geom_line(aes(x = cum_population, 
                y = cum_income, 
                group = year, 
                color = as.factor(year))) +
  theme_minimal()+
  theme(legend.position=c(0.10,0.60), 
        legend.title=element_blank())

```

## Question 6

### Functions 

```{r}

gini<-function(){

gini_coefficient <- lorenz() %>%
  group_by(year) %>% 
  summarise(gini = 1 - 2 * integrate(Vectorize(approxfun(cum_population, cum_income)), 0, 1)$value) 

gini_coefficient

}

```

### Printing Results

```{r}

gini() %>% 
  head(10) %>% 
  knitr::kable("html") %>%
  kable_styling("striped", full_width = FALSE) 


```

### Graphing Results

```{r}

gini() %>% 
  ggplot(aes(x=year,y=gini))+
  geom_line()+
  geom_point()+
  theme_minimal()

```
